#Practical Machine Learning - Prediction Assignment
###Author: Dominic Chan
###date: July 17, 2016



## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Executive Summary

####  How the model was built
The training data was split into 60% training and 40% for testing.  Decision Tree, GBM and Random forest models were considered.  The final model was the Random Forest as it yield the best results.


####  Cross Validation
Cross Validation was done for each model, with K=3


#### Expected out of Sample Error
The random forest model performed the best with model accuracy of 0.9988529


#### Predictions for the 20 test cases
1     B
2     A
3     B
4     A
5     A
6     E
7     D
8     B
9     A
10    A
11    B
12    C
13    B
14    A
15    E
16    E
17    A
18    B
19    B
20    B




## I. Loading the data
- Partition the training data into two 60% training partition and 40% testing partition
```{r}
library(caret)

set.seed(12345)

training_data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",  na.strings=c("NA","#DIV/0!",""))
testing_data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",  na.strings=c("NA","#DIV/0!",""))

inTrain <- createDataPartition(training_data$classe, p=0.6, list=FALSE)
training <- training_data[inTrain, ]
testing <- training_data[-inTrain, ]
```


## II.  Data Prep
- Removed the variables for the record id 'X', username, timestamp as they have no predictive usefulness for this application
- Removed columns that have more than 20% missing values
- Removed columns that are near zero variance

```{r}
testing <- testing[, colMeans(is.na(training)) <= .20]
training <- training[, colMeans(is.na(training)) <= .20]

nzv <- nearZeroVar(training, saveMetrics=FALSE, names=TRUE)
testing <- subset(testing, select=-c(X,user_name,cvtd_timestamp,new_window))
training <- subset(training, select=-c(X,user_name,cvtd_timestamp,new_window))
```

## III.  Model Building
1. Decision Tree (DT)
2. Random Forest (RF)
3. Gradient Boosting Machine (GBM)

### 1. Decision Tree
```{r,warning=FALSE, message=FALSE}
fitControl <- trainControl(method='cv', number = 3)

dt.fit <- train(classe ~ ., method="rpart", trControl=fitControl, data=training)
dt.predict <- predict(dt.fit, testing)
confusionMatrix(dt.predict, testing$classe)
```

### 2. Random Forest
```{r,warning=FALSE, message=FALSE}
rf.fit <- train(classe ~ ., method="rf", trControl=fitControl, data=training, ntree=100)
rf.predict <- predict(rf.fit, testing)
confusionMatrix(rf.predict, testing$classe)
```

### 3. Gradient Boosting Machine
```{r,warning=FALSE, message=FALSE}
gbm.fit <- train(classe ~ ., method="gbm", trControl=fitControl, data=training, verbose = FALSE)
gbm.predict <- predict(gbm.fit, testing)
confusionMatrix(gbm.predict, testing$classe)
```

## IV.  Prediction
```{r,warning=FALSE, message=FALSE}
pred <- predict(rf.fit, newdata=testing_data)
data.frame(pred)
```